---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
  )
```

# evalITR

<!-- badges: start -->
<!-- badges: end -->

## Installation

You can install the development version of evalITR from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("MichaelLLi/evalITR")
```


(Optional) if you have multiple cores, we recommendate using multisession futures and processing in parallel. This would increase computation efficiency and reduce the time to fit the model. 

```r
library(furrr)
library(future.apply)

nworkers <- 4
plan(multisession, workers =nworkers)
```



## Example under sample splitting


This is an example using the `star` dataset (for more information about the dataset, please use `?star`). 

We start with a simple example with one outcome variable (writing scores) and one machine learning algorithm (causal forest). Then we move to incoporate multiple outcomes and compare model performances with several machine learning algorithms. 


To begin, we load the dataset and specify the outcome variable and covariates to be used in the model. Next, we utilize a random forest algorithm to develop an Individualized Treatment Rule (ITR) for estimating the varied impacts of small class sizes on students' writing scores.  Since the treatment is often costly for most policy programs, we consider a case with 20% budget constraint (`plim` = 0.2). The model will identify the top 20% of units who benefit from the treatment most and assign them to with the treatment. We train the model through sample splitting, with the `ratio` between the train and test sets determined by the ratio argument. Specifically, we allocate 70% of the data to train the model, while the remaining 30% is used as testing data (ratio = 0.7).


```{r estimate, message = FALSE}
library(tidyverse)
library(evalITR)

load("data/star.rda")

# specifying the outcome
outcomes <- "g3tlangss"

# specifying the treatment
treatment <- "treatment"

# specifying covariates
covariates <-  star %>% dplyr::select(-c("g3tlangss",
                "g3treadss","g3tmathss","treatment")) %>% 
                colnames()

# estimate ITR 
fit <- run_itr(outcome = outcomes,
               treatment = treatment,
               covariates = covariates,
               data = star,
               algorithms = c("causal_forest"),
               plim = 0.2,
               ratio = 0.7)

# evaluate ITR 
est <- estimate_itr(fit)
```

The`summary()` function displays the following summary statistics: (1) population average prescriptive effect `PAPE`; (2) population average prescriptive effect with a budget constraint `PAPEp`; (3) population average prescriptive effect difference with a budget constraint `PAPDp`. This quantity will be computed with more than 2 machine learning algorithms); (4) and area under the prescriptive effect curve `AUPEC`. For more information about these evaluation metrics, please refer to [Imai and Li (2021)](https://arxiv.org/abs/1905.05389); (5) Grouped Average Treatment Effects `GATEs`.  The details of the methods for this design are given in [Imai and Li (2021)](https://arxiv.org/abs/2203.14511).


```{r sp_summary}
# summarize estimates
summary(est)
```


We plot the estimated Area Under the Prescriptive Effect Curve for the writing score across a range of budget constraints for causal forest.

```{r sp_plot, fig.width=4, fig.height=3,fig.align = "center"}
# plot the AUPEC 
plot(est)

```

## Example under cross-validation

The package also allows estimate ITR with k-folds cross-validation. Instead of specifying the `ratio` argument, we choose the number of folds (`n_folds`). The following code presents an example of estimating ITR with 3 folds cross-validation. In practice, we recommend using 10 folds to get a more stable model performance. 

```{r cv_estimate, message = FALSE}
# estimate ITR 
fit_cv <- run_itr(outcome = outcomes,
               treatment = treatment,
               covariates = covariates,
               data = star,
               algorithms = c("causal_forest"),
               plim = 0.2,
               n_folds = 3)

# evaluate ITR 
est_cv <- estimate_itr(fit_cv)

```

We present the results with 3-folds cross validation and plot the AUPEC. 

```{r cv_summary}
# summarize estimates
summary(est_cv)
```


```{r sv_plot, fig.width=4, fig.height=3,fig.align = "center"}
# plot the AUPEC 
plot(est_cv)
```


## Example with multiple ML algorithms/outcomes

We can estimate ITR with various machine learning algorithms and then compare the performance of each model. The package includes 8 different ML algorithms (causal forest, BART, lasso, boosting trees, random forest, CART, bagging trees, svm).

The package also allows estimate heterogeneous treatment effects on the individual and group-level. On the individual-level, the summary statistics and the AUPEC plot show whether assigning individualized treatment rules may outperform complete random experiment. On the group-level, we specify the number of groups through `ngates` and estimating heterogeneous treatment effects across groups. 

If the original experiment has diverse outcome measures, we develop ITRs for each outcome and use them to estimate the heterogeneous effects across the different outcomes.


```{r multiple, message = FALSE}
# specifying outcomes
outcomes <- c("g3tlangss","g3treadss","g3tmathss")

# specifying covariates
covariates <-  star %>% dplyr::select(-c("g3tlangss","g3treadss","g3tmathss","treatment")) %>% colnames()

# train the model
fit_cv <- run_itr(outcome = outcomes,
               treatment = "treatment",
               covariates = covariates,
               data = star,
               algorithms = c(
                  "causal_forest", 
                  # "bartc",
                  # "svm",
                  "lasso",
                  "boost", 
                  "random_forest",
                  "cart",
                  "bagging"),
               plim = 0.2,
               n_folds = 3)
```

```{r multiple_estimates}
# compute estimates
est_cv <- estimate_itr(fit_cv)
```

```{r multiple_summary}
# summarize estimates
summary(est_cv, outcome = "g3tlangss")
```

We plot the estimated Area Under the Prescriptive Effect Curve for the writing score across different ML algorithms.


```{r multiple_plot, fig.width=8, fig.height=6,fig.align = "center"}
# plot the AUPEC with different ML algorithms
plot(est_cv, outcome = "g3tlangss")


```