---
title: "User Defined ITR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{user_itr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "../man/figures/README-"
  )

library(dplyr)
library(evalITR)

load("../data/star.rda")

# specifying the outcome
outcomes <- "g3tlangss"

# specifying the treatment
treatment <- "treatment"

# specifying the data (remove other outcomes)
star_data <- star %>% dplyr::select(-c(g3treadss,g3tmathss))

# specifying the formula
user_formula <- as.formula(
  "g3tlangss ~ treatment + gender + race + birthmonth + 
  birthyear + SCHLURBN + GRDRANGE + GKENRMNT + GKFRLNCH + 
  GKBUSED + GKWHITE ")

```

Instead of using the ITRs estimated by `evalITR` models, we can define our own ITR and evaluate its performance using the `evaluate_itr` function. The function takes the following arguments:


| Argument | Description              |
|:-------- | :------------------------|
| `user_itr` | a function defined by users that returns a unit-level continuous score for treatment assignment (we assume those that have score less than 0 should not have treatment) |
| `data`   | a data frame   |
| `treatment` | a character string specifying the treatment variable in the `data` |
| `outcome` | a character string specifying the outcome variable in the `data` |
| `budget` | a numeric value specifying the maximum percentage of population that can be treated under the budget constraint |


The function returns an object that contains the estimated GATE, ATE, and AUPEC for the user defined ITR.

```{r user_itr_summary, warning = FALSE, message = FALSE}
# user's own ITR
score_function <- function(data){

  data %>% 
    mutate(score = case_when(
      SCHLURBN == 1 ~ 0.1, # inner-city
      SCHLURBN == 2 ~ 0.2, # suburban
      SCHLURBN == 3 ~ 0.4, # rural
      SCHLURBN == 4 ~ 0.3, # urban
    )) %>%
    pull(score) -> score
    
  return(score)
}

# evalutate ITR
user_itr <- evaluate_itr(
  user_itr = score_function,
  data = star_data,
  treatment = treatment,
  outcome = outcomes,
  budget = 0.2)

# summarize estimates
summary(user_itr)
```


We can extract estimates from the `est` object. The following code shows how to extract the GATE estimates for the writing score with the causal forest algorithm. 

```{r user_itr_gate, warning = FALSE, message = FALSE, fig.width = 6, fig.height = 4}
# plot GATE estimates
library(ggplot2)
summary(user_itr)$GATE %>%
  mutate(group = forcats::as_factor(group)) %>%
  ggplot(., aes(
    x = group, y = estimate,
    ymin = lower , ymax = upper, color = algorithm)) +
  ggdist::geom_pointinterval(
    width = 0.5,
    position = position_dodge(0.5),
    interval_size_range = c(0.8, 1.5),
    fatten_point = 2.5) +  
  theme_bw() +
  theme(panel.grid = element_blank(),
        panel.background = element_blank()) +
  labs(x = "Group", y = "GATE estimate") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "#4e4e4e") +
  scale_color_manual(values = c("#0072B2", "#E69F00", "#56B4E9", "#009E73", "#076f00"))
```  


We plot the estimated Area Under the Prescriptive Effect Curve (AUPEC) for the writing score across a range of budget constraints for user defined ITR.

```{r user_itr_aupec, fig.width = 6, fig.height = 4}
# plot the AUPEC 
plot(user_itr)
```
